Code used to generate paper results : Database comparison and CT characteristics

### 2023-10-24 Paper_code-Database_CT.Rmd
Change in function figures

```{r GO_ORA}

msigdbr(species = "Homo sapiens" , category = "C5") %>% 
  filter(gene_symbol %in% CT_genes$external_gene_name) %>% 
  pull(gene_symbol) %>% 
  unique() %>% 
  length()

@@

as_tibble(go_ora) %>% 
  arrange(desc(Count)) %>% 
  head(12) %>% 
  mutate(Ratio = Count/183) %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  theme(axis.text.y = element_blank(),
        legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title = element_text(size = 10, color = "gray10"))+ 
  geom_text(aes(0, y = Description, label = Description),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 4)

>>>>

All CT genes function

```{r GO_ORA}
msigdbr(species = "Homo sapiens" , category = "C5") %>% 
  filter(gene_symbol %in% CT_genes$external_gene_name) %>% 
  pull(gene_symbol) %>% 
  unique() %>% 
  length()

@@

as_tibble(go_ora) %>% 
  arrange(desc(Count)) %>% 
  head(12) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/177,
                           ONTOLOGY == "CC"~ Count/197,
                           ONTOLOGY == "MF"~ Count/186)) %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  theme(axis.text.y = element_blank(),
        legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title = element_text(size = 10, color = "gray10"))+ 
  geom_text(aes(0, y = Description, label = Description),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 4)

ora_to_plot <- as_tibble(simplify(go_ora))

# Combine the two cilium movement from BP
ora_to_plot[ora_to_plot$ID == "GO:0003341", "Count"] <- 
  ora_to_plot[ora_to_plot$ID == "GO:0003341", "Count"] +
  ora_to_plot[ora_to_plot$ID == "GO:0060294", "Count"]

ora_to_plot <- ora_to_plot %>% 
  arrange(desc(Count)) %>% 
  head(11) %>% 
  filter(ID != "GO:0060294") %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/177,
                           ONTOLOGY == "CC"~ Count/197,
                           ONTOLOGY == "MF"~ Count/186))

ora_to_plot %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  theme(axis.text.y = element_blank(),
        legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title = element_text(size = 10, color = "gray10"))+ 
  geom_text(aes(0, y = Description, label = Description),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 3.7)
```

Is there a difference between meth reg or not ?

```{r GO_ORA_meth}

go_ora_meth <- enrichGO(gene = 
                          filter(CT_genes, regulated_by_methylation)$ensembl_gene_id,
                        keyType = "ENSEMBL",
                        OrgDb = org.Hs.eg.db,
                        ont = "all",
                        readable = TRUE)
go_ora_meth <- simplify(go_ora_meth)

go_ora_not_meth <- enrichGO(gene = 
                          filter(CT_genes, !regulated_by_methylation)$ensembl_gene_id,
                        keyType = "ENSEMBL",
                        OrgDb = org.Hs.eg.db,
                        ont = "all",
                        readable = TRUE)

go_ora_not_meth <- simplify(go_ora_not_meth)

go_ora_meth_plot <- as_tibble(go_ora_meth) %>% 
  arrange(desc(Count)) %>% 
  head(11) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/101,
                           ONTOLOGY == "MF"~ Count/107)) %>%  
  mutate(regulation = "Methylation")

go_ora_not_meth_plot <- as_tibble(go_ora_not_meth) %>% 
  arrange(desc(Count)) %>% 
  head(11) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/76,
                           ONTOLOGY == "CC"~ Count/93))%>%  
  mutate(regulation = "Not methylation")

rbind(go_ora_meth_plot, go_ora_not_meth_plot) %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  facet_wrap(~ regulation) + 
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(size = 6, color = "gray10"),
        axis.title.x = element_text(size = 10, color = "gray10"),
        axis.title.y = element_blank())+ 
  geom_text(aes(0, y = Description, label = ID),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 2)
